---
title: "Los Angeles Crime Statistics"
output: html_document
date: "2024-03-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

**Zane, Cameron, Josh, Nicholas, Niraj**

For our project, we retrieved a data set from https://catalog.data.gov/dataset/crime-data-from-2020-to-present.

This data set reflects incidents of crime in the city of Los Angeles dating back to 2020. There were originally over 918,000 rows, which we have brought down to 200,000 to make it easier to work with. This is still a very large data set. A description of the dataset and each column can be found at https://data.lacity.org/Public-Safety/Crime-Data-from-2020-to-Present/2nrs-mtv8/about_data.

```{r max.print=10}
library(dplyr)
library(lubridate)

crimes <- read.csv("https://raw.githubusercontent.com/ZaneE5/ds202finalproject/main/crimes.csv")

crimes %>% head()
```

**Our questions:**

- What is the most likely victim demographic based on crime committed? 

- Is there a significant change in type of crime committed based on time or location?

- What is the relationship between the date the crime occurred the date it was reported?

- What kinds of weapons are used the most and in what crimes?


To clean the data, we will remove unecessary columns, fix/remove the data that is incomplete/missing, and format the columns to make them easier to work with.

```{r}
#data cleaning steps here

#Columns DR_NO is meaningless to us
#Columns Part.1.2, Premis.Cd, Weapon.Used.Cd, and Mocodes could be useful but we do not know what these columns/codes mean, even with the column descriptions.
#Crm.Cd.1 is the same as Crm.Cd, and we are not interested in Crm.Cd.2-4
#So we remove all these columns to reduce the data to only what we care about

crimes <- select(crimes, c("Date.Rptd", "DATE.OCC", "TIME.OCC", "AREA", "AREA.NAME", "Rpt.Dist.No", "Crm.Cd", "Crm.Cd.Desc", "Vict.Age", "Vict.Sex", "Vict.Descent", "Premis.Desc", "Weapon.Desc", "Status", "Status.Desc", "LOCATION", "Cross.Street", "LAT", "LON"))

#We change some of the chr columns into factors for ease of grouping later
crimes$AREA.NAME <- factor(crimes$AREA.NAME, unique(crimes[,'AREA.NAME']))
crimes$Crm.Cd.Desc <- factor(crimes$Crm.Cd.Desc, unique(crimes[,'Crm.Cd.Desc']))
crimes$Vict.Sex <- factor(crimes$Vict.Sex, unique(crimes[,'Vict.Sex']))
crimes$Vict.Descent <- factor(crimes$Vict.Descent, unique(crimes[,'Vict.Descent']))
crimes$Premis.Desc <- factor(crimes$Premis.Desc, unique(crimes[,'Premis.Desc']))
crimes$Weapon.Desc <- factor(crimes$Weapon.Desc, unique(crimes[,'Weapon.Desc']))
crimes$Status <- factor(crimes$Status, unique(crimes[,'Status']))
crimes$Status.Desc <- factor(crimes$Status.Desc, unique(crimes[,'Status.Desc']))
crimes$LOCATION <- factor(crimes$LOCATION, unique(crimes[,'LOCATION']))
crimes$Cross.Street <- factor(crimes$Cross.Street, unique(crimes[,'Cross.Street']))

#We do a little cleaning on the dates to remove the unecessary 0:00 times
crimes$DATE.OCC <- mdy_hm(crimes$DATE.OCC, tz = "America/Los_Angeles")
crimes$Date.Rptd <- mdy_hm(crimes$Date.Rptd, tz = "America/Los_Angeles")

```


After cleaning, we want to get some of the basic information about the dataset.

```{r}
#marginal summaries
dim(crimes)

str(crimes)

head(crimes)

```
```{r}
library(ggplot2)
filter(crimes, `Crm.Cd.Desc`=="BURGLARY" &! `Weapon.Desc` == "") %>%
ggplot( aes(y=`Weapon.Desc`)) +
  geom_bar()

filter(crimes, `Crm.Cd.Desc`=="VIOLATION OF RESTRAINING ORDER" &! `Weapon.Desc` == "") %>%
ggplot( aes(y=`Weapon.Desc`)) +
  geom_bar()

crimes %>% group_by(`Crm.Cd.Desc`) %>%
  summarize()
```

